Understanding domestic abuse

### 2019-07-29 JeS_fields.tex
Anna update

2) To improve our understanding of the costs of domestic abuse, by quantifying the causal effect of victimisation on a variety of socio-economic outcomes. Current cost estimates by the government do not take into account how educational attainment, income and employment status of victims are affected by victimisation and witnessing abuse as a child \cite{costs}. We will improve the precision of this calculation with the first, comprehensive UK-based exploration of the long-term effects of domestic abuse victimisation.

3) To translate our research into impact by delivering applicable insights for the relevant agencies, including the police and policy-makers. Our research will provide the police with important information to help the optimal geographical and temporal allocation of resources and improving recording practices. In addition, these insights will also help the government to design more effective intervention campaigns, and will be of interest to a wide range of academics.

4) To utilise the rich information contained in the Crime Survey for England and Wales (CSEW) to explore the predictors and consequences of domestic abuse victimisation. In spite of the detailed information contained in the CSEW about respondents, we are not aware of any non-exploratory study using the CSEW to explore the predictors and consequences of domestic abuse victimisation in England and Wales. We aim to take full advantage of the wealth of information in the CSEW in understanding domestic abuse.

@@

Concern about domestic abuse is widespread, and it is rightly identified as a serious societal issue affecting the lives of people from all walks of life. However, the true extent of the problem is still often underestimated, and hard to comprehend: 2 million adults in England and Wales (7.9\% of women and 4.2\% of men) reported to have experienced some form of domestic abuse in the year ending 2018 March \cite{ONS}, and 41\% of all women murdered in England and Wales in this period were victims of domestic homicide. The long-lasting, damaging effects of domestic abuse are not limited to the direct target of abuse, but also has serious developmental implications for children witnessing the abuse. Illustrating the extent of the societal harm caused by domestic abuse, the economic cost of domestic abuse in England and Wales between April, 2016 and March, 2017 was estimated to be as high as \pounds 66 billion \cite{costs}.

@@

Our project fills this gap in our current understanding of domestic abuse by taking advantage of the wealth of information contained in the CSEW, and drawing on other data sources to complement the findings. Applying sophisticated statistical techniques tailored to each analysis (including regression, random forest, propensity score matching), we will identify the individual and neighbourhood-level predictors of domestic abuse victimisation, and the factors affecting victims' willingness to contact the police. We will also explore whether police mis-recording of domestic abuse-related crimes disproportionately affects certain victim groups. Furthermore, we will quantify the short- and long-term causal effects of witnessing domestic abuse on behavioural outcomes in young adulthood (with particular focus on the propensity to engage in violent criminal behaviours) and socio-economic and health-related outcomes in later life. Finally, we will explore how alcohol consumption, spending on gambling, and changes in benefit receipts affect the prevalence of domestic abuse.

>>>>

2) To improve our understanding of the economic costs of domestic abuse, by quantifying the causal effect of victimisation on a variety of socio-economic outcomes. Current cost estimates by the government do not take into account how educational attainment, income and employment status of victims are affected by victimisation and witnessing abuse as a child \cite{costs}. We will improve the precision of this calculation with the first, comprehensive UK-based exploration of the long-term effects of domestic abuse victimisation.

3) To translate our research into impact by delivering applicable insights for the relevant agencies, including the police and policy-makers. Our research will provide the police with important information to facilitate the optimal geographical and temporal allocation of resources and improvement of recording practices. In addition, these insights will also help the government to design more effective intervention campaigns to protect victims, and will be of interest to a wide range of academics.

4) To exploit the rich information contained in the Crime Survey for England and Wales (CSEW) to explore the predictors and consequences of domestic abuse victimisation. Despite the rich, detailed information contained in the CSEW about respondents, we are not aware of any non-exploratory study using the CSEW to understand the predictors and consequences of domestic abuse victimisation in England and Wales. We aim to take full advantage of the wealth of information in the CSEW in understanding domestic abuse.

@@

Concern about domestic abuse is widespread, and it is rightly identified as a serious societal issue affecting the lives of people from all walks of life. However, the true extent of the problem is still often underestimated, and hard to comprehend: 2 million adults in England and Wales (7.9\% of women and 4.2\% of men) reported to have experienced some form of domestic abuse in the year ending 2018 March \cite{ONS}, and 41\% of all women murdered in England and Wales in this period were victims of domestic homicide. The long-lasting, damaging effects of domestic abuse are not limited to the direct target of abuse, but also has serious developmental implications for children witnessing the abuse. Illustrating the extent of the societal harm caused by domestic abuse, the economic cost of domestic abuse in England and Wales was estimated to be as high as \pounds 66 billion in the year ending March, 2017 \cite{costs}.

@@

Our project fills this gap in our current understanding of domestic abuse, by taking advantage of the wealth of information contained in the CSEW, and drawing on other data sources to complement the findings. Applying sophisticated statistical techniques tailored to each analysis (including regression, random forest, and propensity score matching), we will identify the individual and neighbourhood-level predictors of domestic abuse victimisation, and the factors affecting victims' willingness to contact the police. We will also explore whether police mis-recording of domestic abuse-related crimes disproportionately affects certain victim groups. Furthermore, we will quantify the short- and long-term causal effects of witnessing domestic abuse on behavioural outcomes in young adulthood (with particular focus on the propensity to engage in violent criminal behaviours) and socio-economic and health-related outcomes in later life. Finally, we will explore how alcohol consumption, spending on gambling, and changes in benefit receipts affect the prevalence of domestic abuse.
